Paper To-Do

### 2016-02-14 Rmd/empirical-yields.Rmd
Many updates. Up through new analysis.

```{r, eval = FALSE}
library(agroEcoTradeoff)
library(matrixStats)
library(doMC)
library(mgcv)

p_root <- lmisc::set_base_path()
p_clim <- fp(p_root, "external/input_devel/climate")
p_yield <- fp(p_root, "external/input_devel/yield")
p_sgrids <- fp(p_root, "external/input_devel/soil") 
p_basedat <- paste0("/Users/lestes/Dropbox/projects/agroEcoTradeoff/external/",
                    "base_data/")
p_cropland <- full_path(p_root, "external/input_devel/cropland")

@@

sgam1 <- gam(YLDS ~ s(MAP, k = 3) + s(TMU, k = 3)  + s(SLHW, k = 3) + 
               s(SLCL, k = 3) + s(SLOC, k = 3) + s(SSRMIII, k = 3), 
             family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam2 <- gam(YLDS ~ poly(MAP, 2) + poly(TMU, 2)  + poly(SLHW, 2) + 
               poly(SLCL, 2) + poly(SLOC, 2) + poly(SSRMIII, 2), 
             family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam3 <- gam(YLDS ~ poly(DJFMP, 2) + poly(TMU, 2)  + poly(SLHW, 2) + 
               poly(SLCL, 2) + poly(SLOC, 2) + poly(SSRMIII, 2), 
             family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam4 <- gam(YLDS ~ poly(JFMP, 2) + poly(TMU, 2)  + poly(SLHW, 2) + 
               poly(SLCL, 2) + poly(SLOC, 2) + poly(SSRMIII, 2), 
             family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam5 <- gam(YLDS ~ poly(JFMP, 2) + poly(TMU, 2)  + poly(SLHW, 2) + 
               poly(SLCL, 2) + poly(SLOC, 2) + poly(SSRMIII, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam6 <- gam(YLDS ~ poly(JFMP, 2) + poly(TMU, 2)  + poly(SLHW, 2) + 
               poly(SLCL, 2) + poly(SSRMIII, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam7 <- gam(YLDS ~ poly(JFMP, 2) + poly(GDD, 2) + poly(SLHW, 2) + 
               poly(SLCL, 2) + poly(SSRMIII, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam8 <- gam(YLDS ~ poly(JFMP, 2) + poly(STMX, 2) + poly(SLHW, 2) + 
               poly(SLCL, 2) + poly(SSRMIII, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam9 <- gam(YLDS ~ poly(JFMP, 2) + poly(NKJ, 2) + poly(SLHW, 2) + 
               poly(SLCL, 2) + poly(SSRMIII, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam10 <- gam(YLDS ~ poly(JFMP, 2) + poly(NKF, 2) + poly(SLHW, 2) + 
               poly(SLCL, 2) + poly(SSRMIII, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam11 <- gam(YLDS ~ poly(JFMP, 2) + poly(ATMX, 2) + poly(SLHW, 2) + 
               poly(SLCL, 2) + poly(SSRMIII, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam12 <- gam(YLDS ~ poly(JFMP, 2) + poly(TMU, 2) + poly(SLHW, 2) + 
               poly(SLCL, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam13 <- gam(YLDS ~ poly(JFMP, 2) + poly(GDD, 2) + KDD + poly(SLHW, 2) + 
               poly(SLCL, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam14 <- gam(YLDS ~ poly(JFMP, 2) + poly(TMU, 2) + poly(SLHW, 2) + 
               poly(SLCL, 2) + poly(SSRMIII, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam15 <- gam(YLDS ~ poly(JFMP, 2) + poly(GDD, 2) + poly(SLHW, 2) + 
               poly(SLCL, 2) + poly(SSRMIII, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
            data = regnew[CROP == "soy"])
sgam16 <- gam(YLDS ~ poly(JFMP, 2) + poly(TMU, 2) + poly(SLHW, 2) + 
               poly(SLCL, 2) + poly(SSRMIII, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
            data = regnew[CROP == "soy"])

summary(sgam1)
plot.gam(sgam1, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam2)  #+ poly version
plot.gam(sgam2, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam3) #+ DJFMP 
plot.gam(sgam3, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam4) #+ Weakly higher for JFMP
plot.gam(sgam4, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam5) #+ s(x,y)
plot.gam(sgam5, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam6) #- dropping SLOC doesn't lose too much, wasn't sensible
plot.gam(sgam6, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam7) # <-> swapping in GDD doesn't change much
plot.gam(sgam7, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam8) #+ STMX 
plot.gam(sgam8, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam9) #- NKJ loses a bit, but looks more sensible
plot.gam(sgam9, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam10) #<-> NKF loses a bit, but wrong direction
plot.gam(sgam10, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam11) #<-> Annual TMx, same deal
plot.gam(sgam11, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam12) #<-> TMU still increasing, even with SSRMIII dropped
plot.gam(sgam12, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam13) #<-> KDD added, going in wrong direction
plot.gam(sgam13, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam14) #<-> same as 6
plot.gam(sgam14, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam15) #<-> Full data version of 7 (with GDD), becomes significant
plot.gam(sgam15, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam16) #<-> Full data version of 14 (Tmu)
plot.gam(sgam16, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)

@@

# gams to keep 
save(gam13, gam14, gam20, gam21, gam23, sgam7, sgam14, sgam15, sgam16, 
     file = fp(p_yield, "empirical/gams1.rda"))

@@

yldralbmsk <- raster::mask(yldralb, mgrid)
writeRaster(yldralbmsk, 
            filename = fp(p_yield, "empirical/yield_surfaces2_albmsk.tif"))

@@

# Bruinsma 2007-2050 growth
bruinsma <- c(maize = 3224 / 1674, soy = 2124 / 1257)

# adjusted DSSAT factors
potmus <- curr_mu * c(bruinsma)  # 4.43 maize, 3.4 soy
potmu_mz <- potmus[1] / (dssat_cmus[[1]] / 1000)  # maize on cropland mean
potmu_soy <- potmus[2] / (dssat_mus[[2]] / 1000)  # soy on overall mean

@@

# plot(soyfao[[1]])
mzfao <- brick(stack(mzfao))
soyfao <- brick(stack(soyfao))
writeRaster(mzfao, filename = fp(p_yield, "empirical/maize-faoadj.tif"))
writeRaster(soyfao, filename = fp(p_yield, "empirical/soy-faoadj.tif"))

>>>>

```{r, eval = FALSE}
library(agroEcoTradeoff)
library(matrixStats)
library(doMC)
library(mgcv)
library(rworldxtra)

setwd("~/Dropbox/publications/zambiaToff/private/zambiaToff/")
p_root <- lmisc::set_base_path()
p_clim <- fp(p_root, "external/input_devel/climate")
p_yield <- fp(p_root, "external/input_devel/yield")
p_sgrids <- fp(p_root, "external/input_devel/soil") 
p_basedat <- paste0("/Users/lestes/Dropbox/projects/agroEcoTradeoff/external/",
                    "base_data/")
p_cropland <- full_path(p_root, "external/input_devel/cropland")
p_figures <- full_path(p_root, "paper/figures/")

@@

sgam1 <- gam(YLDS ~ s(MAP, k = 3) + s(TMU, k = 3)  + s(SLHW, k = 3) + 
               s(SLCL, k = 3) + s(SLOC, k = 3) + s(SSRMIII, k = 3), 
             family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam2 <- gam(YLDS ~ poly(MAP, 2) + poly(TMU, 2)  + poly(SLHW, 2) + 
               poly(SLCL, 2) + poly(SLOC, 2) + poly(SSRMIII, 2), 
             family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam3 <- gam(YLDS ~ poly(DJFMP, 2) + poly(TMU, 2)  + poly(SLHW, 2) + 
               poly(SLCL, 2) + poly(SLOC, 2) + poly(SSRMIII, 2), 
             family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam4 <- gam(YLDS ~ poly(JFMP, 2) + poly(TMU, 2)  + poly(SLHW, 2) + 
               poly(SLCL, 2) + poly(SLOC, 2) + poly(SSRMIII, 2), 
             family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam5 <- gam(YLDS ~ poly(JFMP, 2) + poly(TMU, 2)  + poly(SLHW, 2) + 
               poly(SLCL, 2) + poly(SLOC, 2) + poly(SSRMIII, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam6 <- gam(YLDS ~ poly(JFMP, 2) + poly(TMU, 2)  + poly(SLHW, 2) + 
               poly(SLCL, 2) + poly(SSRMIII, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam7 <- gam(YLDS ~ poly(JFMP, 2) + poly(GDD, 2) + poly(SLHW, 2) + 
               poly(SLCL, 2) + poly(SSRMIII, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam8 <- gam(YLDS ~ poly(JFMP, 2) + poly(STMX, 2) + poly(SLHW, 2) + 
               poly(SLCL, 2) + poly(SSRMIII, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam9 <- gam(YLDS ~ poly(JFMP, 2) + poly(NKJ, 2) + poly(SLHW, 2) + 
               poly(SLCL, 2) + poly(SSRMIII, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam10 <- gam(YLDS ~ poly(JFMP, 2) + poly(NKF, 2) + poly(SLHW, 2) + 
               poly(SLCL, 2) + poly(SSRMIII, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam11 <- gam(YLDS ~ poly(JFMP, 2) + poly(ATMX, 2) + poly(SLHW, 2) + 
               poly(SLCL, 2) + poly(SSRMIII, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam12 <- gam(YLDS ~ poly(JFMP, 2) + poly(TMU, 2) + poly(SLHW, 2) + 
               poly(SLCL, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam13 <- gam(YLDS ~ poly(JFMP, 2) + poly(GDD, 2) + KDD + poly(SLHW, 2) + 
               poly(SLCL, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam14 <- gam(YLDS ~ poly(JFMP, 2) + poly(TMU, 2) + poly(SLHW, 2) + 
               poly(SLCL, 2) + poly(SSRMIII, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
# sgam14a <- gam(YLDS ~ poly(MAP, 2) + poly(TMU, 2) + poly(SLHW, 2) + 
#                poly(SLCL, 2) + poly(SSRMIII, 2) + s(x, y), 
#              family = gaussian(link = "log"), method = "REML", 
#             data = regreds2[YST == "mu"])
# summary(sgam14a)
# plot.gam(sgam14a, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
sgam15 <- gam(YLDS ~ poly(JFMP, 2) + poly(GDD, 2) + poly(SLHW, 2) + 
               poly(SLCL, 2) + poly(SSRMIII, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
            data = regnew[CROP == "soy"])
sgam16 <- gam(YLDS ~ poly(JFMP, 2) + poly(TMU, 2) + poly(SLHW, 2) + 
               poly(SLCL, 2) + poly(SSRMIII, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
            data = regnew[CROP == "soy"])

summary(sgam1)
plot.gam(sgam1, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam2)  #+ poly version
plot.gam(sgam2, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam3) #+ DJFMP 
plot.gam(sgam3, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam4) #+ Weakly higher for JFMP
plot.gam(sgam4, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam5) #+ s(x,y)
plot.gam(sgam5, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam6) #- dropping SLOC doesn't lose too much, wasn't sensible
plot.gam(sgam6, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam7) # <-> swapping in GDD doesn't change much
plot.gam(sgam7, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam8) #+ STMX 
plot.gam(sgam8, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam9) #- NKJ loses a bit, but looks more sensible
plot.gam(sgam9, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam10) #<-> NKF loses a bit, but wrong direction
plot.gam(sgam10, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam11) #<-> Annual TMx, same deal
plot.gam(sgam11, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam12) #<-> TMU still increasing, even with SSRMIII dropped
plot.gam(sgam12, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam13) #<-> KDD added, going in wrong direction
plot.gam(sgam13, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam14a) #<-> same as 6
plot.gam(sgam14a, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam15) #<-> Full data version of 7 (with GDD), becomes significant
plot.gam(sgam15, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam16) #<-> Full data version of 14 (Tmu)
plot.gam(sgam16, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)

@@

# gams to keep 
save(gam13, gam14, gam20, gam21, gam23, sgam7, sgam14, sgam15, sgam16, 
     file = fp(p_yield, "empirical/gams1.rda"))
# load(fp(p_yield, "empirical/gams1.rda"))
summary(gam13)
summary(sgam14)

@@

yldralbmsk <- raster::mask(yldralb, mgrid)
writeRaster(yldralbmsk, 
            filename = fp(p_yield, "empirical/yield_surfaces2_albmsk.tif"))
# yldralbmsk <- brick(fp(p_yield, "empirical/yield_surfaces2_albmsk.tif"))
# names(yldralbmsk) <- c("maizer1", "maizer2", "maizef1", "maizef2", "maizef3",
#                        "soyr1", "soyr2", "soyf1", "soyf2")

@@

# Bruinsma 2007-2050 growth
bruinsma <- c(maize = 3224 / 1674, soy = 2124 / 1257)
fao <- fread(fp(p_yield, "empirical/zambia-maize-soy-yield.csv"))
fao2005 <- fao[year > 2004,]
yrates <- sapply(c("Maize", "Soybeans"), function(i) {
  fao[crop == i, coef(lm(yield / yield[1] ~ year))[2]]
})
ymult <- 1 + yrates * (2050-2014)  # fao yield multipliers

# take average of Bruinsma and FAO projection to account for maize
ymults <- c("maize" = sum(ymult[1] * 0.25, bruinsma[1] * 0.75), bruinsma[2])

# adjusted DSSAT factors
potmus <- curr_mu * c(ymults)  # 4.43 maize, 3.4 soy
potmu_mz <- potmus[1] / (dssat_cmus[[1]] / 1000)  # maize on cropland mean
potmu_soy <- potmus[2] / (dssat_mus[[2]] / 1000)  # soy on overall mean

@@

# plot(soyfao[[1]])
mzfao <- brick(stack(mzfao))
soyfao <- brick(stack(soyfao))
writeRaster(mzfao, filename = fp(p_yield, "empirical/maize-faoadj.tif"), 
            overwrite = TRUE)
writeRaster(soyfao, filename = fp(p_yield, "empirical/soy-faoadj.tif"), 
            overwrite = TRUE)
# mzfao <- brick(fp(p_yield, "empirical/maize-faoadj.tif"))
# soyfao <- brick(fp(p_yield, "empirical/soy-faoadj.tif"))

data("countriesHigh")
zambia <- countriesHigh[countriesHigh@data$SOVEREIGNT == "Zambia", ]
zambia <- spTransform(zambia, mzfao@crs)

png(fp(p_figures, "yields.png"), width = 800, height = 400)
par(mfrow = c(1, 2), mar = c(0, 0, 0, 4))
plot(zambia, col = "grey80", border = "transparent")
plot(mzfao[[1]], add = TRUE)
plot(zambia, col = "grey80", border = "transparent")
plot(soyfao[[1]], add = TRUE)
dev.off()
```

### 2016-02-14 Rmd/paper-todo.Rmd
Many updates. Up through new analysis.

+ Near-term: See if we can bias-correct using NPP.  
+ Longer-term: DSSAT yields -> machine learning (Nate)

@@

+ Near-term: Code in impact based on intactness + protected weight
+ Longer-term: Code in Newbold model (Laura and Tobias)

@@

Analysis/fix | Drop-dead date | Date Finished
------------- | -------------  | -------------
NPP-crop correct  | 29 Jan |
Biodiversity fix  | 31 Jan |
Crop overlaps | ??? |
Yield variability |  ??? |
Combined optimization, back out weights? |  ???

>>>>

+ Near-term: See if we can bias-correct using NPP.  __Not workable__
+ Longer-term: DSSAT yields -> machine learning (Nate) __Done, with GAMs__

@@

+ Near-term: Code in impact based on intactness + protected weight __Done__
+ Longer-term: Code in Newbold model (Laura and Tobias) __Pending__

@@

Analysis/fix | Drop-dead date | Date Finished
------------- | -------------  | -------------
NPP-crop correct  | 29 Jan | Not pursued
Biodiversity fix  | 31 Jan | 11 Feb
Crop overlaps | ??? | 
Yield variability |  ??? | Not pursued
Combined optimization, back out weights? |  ???
